A Pandemic-Scale Ancestral Recombination Graph for SARS-CoV-2

### 2025-10-09 latex/main.tex
Updates from Overleaf

\subsection*{Comparison with RIPPLES}
RIPPLES~\cite{Turakhia2022} is a method for detecting 
recombination events using a tree inferred by UShER. Recombination is 
detected by first identifying plausible branches carrying multiple
mutations (at least 3, by default). For each candidate branch,
RIPPLES then systematically splits the sequence into 2 or 3
segments, and finds parsimonious placements for each segment independently. 
A recombination event at that node is then flagged if the overall parsimony of these multiple 
placements sufficiently improves over the original single-parent
origin in the tree (the ``parsimony improvement'' parameter, $p=3$, by default).
RIVET~\cite{Smith2023tracking} is a tool for quality control
and visualisation of these recombination events.

@@

\subsection*{Recombination intervals}

@@

We find an overall preponderance of breakpoint intervals towards the 3’ end of the genome,
particularly at the left boundary of the Spike gene (Figure~\ref{fig:recombinant_intervals}A,B),
consistent with patterns observed in previous studies
\cite{Bobay2020recombination,Turakhia2022,Smith2023tracking,Li2024CovRecomb}.
While this might result from a higher recombination rate in this region,
it may also be due to the higher levels of polymorphism in the Spike gene,
allowing recombination endpoints to be estimated more precisely.

>>>>

\subsection*{Comparison with RIPPLES}

RIPPLES~\cite{Turakhia2022} is a method for detecting 
recombination events using a tree inferred by UShER.
Recombination is detected by first identifying plausible branches
carrying multiple mutations (at least 3, by default).
For each candidate branch,
RIPPLES then systematically splits the sequence into 2 or 3 segments, and
finds parsimonious placements for each segment independently.
A recombination event at that node is then flagged if
the overall parsimony of these multiple placements
sufficiently improves over the original single-parent origin in the tree
(the ``parsimony improvement'' parameter, $p=3$, by default).
RIVET~\cite{Smith2023tracking} is a tool for quality control
and visualization of these recombination events.

@@

\subsection*{Recombination breakpoint intervals}

@@

We find an overall preponderance of breakpoint intervals
towards the 3’ end of the genome,
particularly at the left boundary of the Spike gene
(Figure~\ref{fig:recombinant_intervals}A,B),
consistent with patterns observed in previous
studies\cite{Bobay2020recombination,Turakhia2022,
Smith2023tracking,Li2024CovRecomb}.
While this might result from a higher recombination rate in this region,
it may also be due to the higher levels of polymorphism in the Spike gene,
allowing recombination endpoints to be estimated more precisely.
